### 2017-06-08 sheets/talk-6/sheet-6.tex
Typo hotfix: s/diagram*/diagram/
Sample solution 6.9

\begin{answer}
  \begin{proof}
    We show that $S := P$ is the coproduct of $X$ and $Y$ together with the projection functions $p_1$ and $p_2$.
    First note that because $Z$ is initial $s$ and $t$ are unique maps.
    Let
    $ \xymatrix{
      X \ar[r]^{f_1} & A & Y \ar[l]_{f_2}
    } $
    be a diagramm in $\cat{A}$.
    We have to show that there is a unique map $\bar{f}' \from P \to A$ shuch that
    \[ \xymatrix{
      X \ar[rdd]_{f_1}\ar[rd]^{p_1} & & Y \ar[ldd]^{f_2} \ar[ld]_{p_2} \\
      & P \ar@{.>}[d]|{\bar{f}'} & \\
      & A &
    } \]
    commutes. We know that the diagramm
    \[ \xymatrix{
      Z \ar[r]_s \ar[d]^t & X \ar[d]_{f_1} \\
      Y \ar[r]^{f_2} & A
    } \]
    commutes because $Z$ is initial.
    Now because this square is then a pushout square there exists a unique map $\bar{f}\from P \to A$ such that
    \[ \xymatrix{
      Z \ar[r]_s \ar[d]^t & X \ar[d]_{p_1} \ar[ddr]^{f_1} & \\
      Y \ar[r]^{p_2} \ar[drr]_{f_2}& P \ar@{.>}[dr]|{\bar{f}} & \\
      & & A  \\
    } \]
    commutes. We see that we can choose $\bar{f}' := \bar{f}$.
    Suppose we have another $\tilde{f} \from P \to A$ such that
    \[ \xymatrix{
      X \ar[rdd]_{f_1}\ar[rd]^{p_1} & & Y \ar[ldd]^{f_2} \ar[ld]_{p_2} \\
      & P \ar@{.>}[d]|{\tilde{f}} & \\
      & A &
    } \]
    commutes. Then
    \[ \xymatrix{
      Z \ar[r]_s \ar[d]^t & X \ar[d]_{p_1} \ar[ddr]^{f_1} & \\
      Y \ar[r]^{p_2} \ar[drr]_{f_2}& P \ar@{.>}[dr]|{\tilde{f}} & \\
      & & A  \\
    } \]
    would also commute (because $S$ is initial) but $\bar{f}$ is unique, hence $\bar{f} = \tilde{f}$.
  \end{proof}

@@

>>>>

\begin{answer}
  \begin{proof}
    We show that $S := P$ is the coproduct of $X$ and $Y$ together with the projection functions $p_1$ and $p_2$.
    First note that because $Z$ is initial $s$ and $t$ are unique maps.
    Let
    $ \xymatrix{
      X \ar[r]^{f_1} & A & Y \ar[l]_{f_2}
    } $
    be a diagram in $\cat{A}$.
    We have to show that there is a unique map $\bar{f}' \from P \to A$ shuch that
    \[ \xymatrix{
      X \ar[rdd]_{f_1}\ar[rd]^{p_1} & & Y \ar[ldd]^{f_2} \ar[ld]_{p_2} \\
      & P \ar@{.>}[d]|{\bar{f}'} & \\
      & A &
    } \]
    commutes. We know that the diagram
    \[ \xymatrix{
      Z \ar[r]_s \ar[d]^t & X \ar[d]_{f_1} \\
      Y \ar[r]^{f_2} & A
    } \]
    commutes because $Z$ is initial.
    Now because this square is then a pushout square there exists a unique map $\bar{f}\from P \to A$ such that
    \[ \xymatrix{
      Z \ar[r]_s \ar[d]^t & X \ar[d]_{p_1} \ar[ddr]^{f_1} & \\
      Y \ar[r]^{p_2} \ar[drr]_{f_2}& P \ar@{.>}[dr]|{\bar{f}} & \\
      & & A  \\
    } \]
    commutes. We see that we can choose $\bar{f}' := \bar{f}$.
    Suppose we have another $\tilde{f} \from P \to A$ such that
    \[ \xymatrix{
      X \ar[rdd]_{f_1}\ar[rd]^{p_1} & & Y \ar[ldd]^{f_2} \ar[ld]_{p_2} \\
      & P \ar@{.>}[d]|{\tilde{f}} & \\
      & A &
    } \]
    commutes. Then
    \[ \xymatrix{
      Z \ar[r]_s \ar[d]^t & X \ar[d]_{p_1} \ar[ddr]^{f_1} & \\
      Y \ar[r]^{p_2} \ar[drr]_{f_2}& P \ar@{.>}[dr]|{\tilde{f}} & \\
      & & A  \\
    } \]
    would also commute (because $S$ is initial) but $\bar{f}$ is unique, hence $\bar{f} = \tilde{f}$.
  \end{proof}

@@

\begin{answer}
  \begin{itemize}
    \item[``$\Rightarrow$'']
      \emph{Counterexample:}
      Let $\cat{A} := \Set, E := \emptyset, X := \nat, Y := \mathbb{B}$.
      Define $f(n) := \text{even}(n)$ and $g(n) := \text{odd}(n)$. $i$ is the (empty) inclusion function.
      Then
      \[ \xymatrix{
        E \ar[r]^i \ar[d]_i &
        X \ar[d]^g \\
        X \ar[r]_f &
        Y
      } \]
      is clearly a pullback because no number is even and odd at the same time.
      Now let $A := \nat$, $f_1(n) := n$ and $f_2(n) := n+1$.
      Then the diagram
      \[ \xymatrix{
        A \ar[r]_{f_2} \ar[d]^{f_1} & Y \ar[d]_{g} \\
        X \ar[r]^f & Z
      } \]
      commutes because $\text{even}(n) \Leftrightarrow \text{odd}(n+1)$.
      But there exists no function $\bar{f} \from \nat \to \emptyset$ at all.
    \item[``$\Leftarrow$'']
      \begin{proof}
        Let
        \[ \xymatrix{
          E \ar[r]^i \ar[d]_i &
          X \ar[d]^g \\
          X \ar[r]_f &
          Y
        } \]
        be the pullback of the diagram
        $ \xymatrix{
          X \ar[r]^f & Y & X\ar[l]_g \\
        } $.
        We show that $\xymatrix@1{E \ar[r]^i &X \ar@<.5ex>[r]^{f} \ar@<-.5ex>[r]_g &Y}$ is the equalizer of $\xymatrix@1{X \ar@<.5ex>[r]^{f} \ar@<-.5ex>[r]_g &Y}$.
        First we note that the diagramm $\xymatrix@1{E \ar[r]^i &X \ar@<.5ex>[r]^{f} \ar@<-.5ex>[r]_g &Y}$ is indeed a fork as it commutes by assumption.
        Let
        $ \xymatrix{
          A \ar[r]^f & X \ar@<.5ex>[r]^s \ar@<-.5ex>[r]_t & Y
        } $
        be a fork. Then
        \[ \xymatrix{
          A \ar[r]_{f} \ar[d]^{f} & X \ar[d]_{t} \\
          X \ar[r]^s & Y
        } \]
        commutes. Therefore exists a unique map $\bar{f}\from A \to P$ such that
        \[ \xymatrix{
          A \ar[drr]^{f} \ar[ddr]_{f} \ar@{.>}[dr]|{\bar{f}} & & \\
          & P \ar[r]_{i} \ar[d]^{i} & X \ar[d]_{t} \\
          & X \ar[r]^s & Y
        } \]
        commutes. We show that this $\bar{f}$ is the only map such that
        \[ \xymatrix{
          A \ar[rd]^f \ar@{.>}[dd]|{\bar{f}} & & \\
          & X \ar@<.5ex>[r]^s \ar@<-.5ex>[r]_t & Y \\
          P \ar[ru]_i & &
        } \]
        commutes. First it is clear that this diagramm commutes for $\bar{f}$. Now let $\tilde{f}$ be another map such that
        \[ \xymatrix{
          A \ar[rd]^f \ar@{.>}[dd]|{\tilde{f}} & & \\
          & X \ar@<.5ex>[r]^s \ar@<-.5ex>[r]_t & Y \\
          P \ar[ru]_i & &
        } \]
        commutes. But then
        \[ \xymatrix{
          A \ar[drr]^{f} \ar[ddr]_{f} \ar@{.>}[dr]|{\tilde{f}} & & \\
          & P \ar[r]_{i} \ar[d]^{i} & X \ar[d]_{t} \\
          & X \ar[r]^s & Y
        } \]
        also commutes. By the assumption that $\bar{f}$ is unique we have $\bar{f} = \tilde{f}$. \qedhere
      \end{proof}
  \end{itemize}
\end{answer}
